Restrictive Clearance Comparison

### 2024-04-05 data-raw/RestrictiveClearance_comparison.Rmd
Clarified some functions, added systematic comparisons between constant and optimized kelim models

```{r library_loading}
devtools::load_all()
library(httk)
my_pk <- readRDS("data-raw/all_cvt_pk_JOINT.rds")
```

@@

```{r running-invivoPKfit}
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% common_chems[1:8]),
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))

@@

new_pk <- do_fit(new_pk, n_cores = 6) 
# Multidplyr error because of forgotten browser() call within log_likelihood
new_pk$fit %>% unnest(fit) %>% View()

coef(new_pk) %>% View()
eval_tkstats(new_pk) %>% View()

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "kelim_compare_2024_0329.png"),
       height = 11,
       width = 8.5)
```

```{r testing_optimx}
test_optimx <- do.call(
          optimx::optimx,
          args = c(
            list(par = opt_params,
                 fn = log_likelihood,
                 lower = lower_params,
                 upper = upper_params),
            #method and control
            settings_optimx,
            #... additional args to log_likelihood
            list(
              const_params = const_params,
              data = data,
              data_sigma_group = data$data_sigma_group,
              modelfun = modelfun,
              dose_norm = dose_norm,
              log10_trans = log10_trans,
              negative = TRUE,
              force_finite = TRUE,
              suppress.messages = FALSE
            ) #end list()
          ) #end args = c()
        ) #end do.call
```

@@

rest_wm <- get_winning_model(rest_pk)
rest_AIC <- AIC(rest_pk)
rest_rsq <- rsq(rest_pk)

eval_tkstats(rest_pk) %>% View()
residuals(rest_pk) %>% View()

pl <- plot(rest_pk, n_interp = 12, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "RestrictiveCL_AllFits_Final_202402_08.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

# Non-restrictive fits (constant kelim)

pl <- plot(nonrest_pk, n_interp = 12, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "Non-restrictiveCL_AllFits_Final_202402_08.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

# try dawson
httk::load_dawson2021()

@@

>>>>

```{r library_loading}
devtools::load_all()
library(httk)
library(tidyverse)
my_pk <- readRDS("data-raw/all_cvt_pk_JOINT.rds")
```

@@

```{r running-invivoPKfit}
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% common_chems),
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))

@@

new_pk <- do_fit(new_pk, n_cores = 10) 
# Multidplyr error because of forgotten browser() call within log_likelihood
new_pk$fit %>% unnest(fit) %>% View()

coef(new_pk) %>% View()
eval_tkstats(new_pk, finite_only = FALSE) %>% View()
get_winning_model(new_pk) %>% View("winning_model")

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "kelim_compare_2024_0329.png"),
       height = 11,
       width = 8.5)

@@

new_wm <- get_winning_model(new_pk)
new_AIC <- AIC(new_pk)
new_rsq <- rsq(new_pk)

new_tkstats <- eval_tkstats(new_pk, finite_only = FALSE)

residuals(new_pk) %>% View()

pl <- plot(new_pk, n_interp = 12, use_scale_conc = TRUE, best_fit = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "BestFits_Final_2024_0405.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

@@

Comparing total clearance estimates for those where a restrictive or non-restrictive model
was the "best fit" as determined by AIC.

```{r cltot-comparison}
new_cltot2 <- new_tkstats %>% 
  # filter(str_detect(model, "rest")) %>%
  select(Chemical, Species, Media, Route, 
         model, method,
         CLtot.tkstats, CLtot.nca) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(clearance_type = ifelse(str_detect(model, "nonrest"),
                                "NonRestrictive",
                                ifelse(str_detect(model, "rest"),
                                       "Restrictive",
                                       "optimized kelim"))) %>%
  rowwise() %>%
  mutate(CLtot.httk_rest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = TRUE),
         CLtot.httk_nonrest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = FALSE))

new_cltot2 %>%
  ggplot(aes(x = Chemical)) +
  geom_point(aes(y = CLtot.httk_rest),shape = 19, size = 2, color = "green4") +
  geom_point(aes(y = CLtot.httk_nonrest),shape = 19, size = 2, color = "orange2") +
  geom_point(aes(y = CLtot.tkstats), shape = 1, size = 3, color = "black", stroke = 1) +
  scale_y_log10(labels = scales::label_math(format = log10)) +
  geom_hline(yintercept = c(0.01, 0.1, 1, 10)) +
  coord_flip() +
  facet_grid(rows = vars(clearance_type),
             scales = "free", space = "free") +
  labs(title = "Total Clearance for constant kelim \"best fits\" (empty)\nand httk calculations (filled)") +
  theme_bw() +
  theme(axis.line = element_line(linewidth = 1),
        axis.text = element_text(size = 12),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "white"))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "Cltot_backcompare_2024_0405.png"),
       height = 4,
       width = 6)

```{r chemical-correlation}
new_AIC %>%
  ungroup() %>%
  filter(model %in% c("model_1comp", "model_2comp"),
         Species %in% "rat") %>%
  group_by(Chemical) %>%
  filter(AIC == min(AIC)) %>%
  mutate(estimate = "nonConst") %>%
  distinct(Chemical, model, estimate) -> nAIC_nonConst
new_AIC %>%
  ungroup() %>%
  filter(model %in% c("model_1comp_rest", "model_2comp_rest"),
         Species %in% "rat") %>%
  group_by(Chemical) %>%
  filter(AIC == min(AIC)) %>%
  mutate(estimate = "Rest") %>%
  distinct(Chemical, model, estimate) -> nAIC_rest
new_AIC %>%
  ungroup() %>%
  filter(model %in% c("model_1comp_nonrest", "model_2comp_nonrest"),
         Species %in% "rat") %>%
  group_by(Chemical) %>%
  filter(AIC == min(AIC)) %>%
  mutate(estimate = "nonRest") %>%
  distinct(Chemical, model, estimate) -> nAIC_nonrest

full_nAIC <- bind_rows(nAIC_nonConst, nAIC_rest) %>%
  bind_rows(nAIC_nonrest)

new_tkstats_full <- get_tkstats(new_pk)

new_tkstats_full %>%
  filter(Species %in% "rat", !(model %in% "model_flat")) %>%
  distinct(Chemical, model, CLtot) %>%
  inner_join(full_nAIC) %>%
  na.omit() %>% 
  distinct(Chemical, estimate, CLtot) %>%
  pivot_wider(names_from = estimate,
              values_from = CLtot) %>%
  rowwise() %>%
  mutate(CLtot.httk_rest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = TRUE),
         CLtot.httk_nonrest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = FALSE)) %>%
  ungroup() %>%
  mutate(A_rest = log(nonConst) - log(CLtot.httk_rest),
         A_nonrest = log(nonConst) - log(CLtot.httk_nonrest),
         B_rest = log(Rest) - log(CLtot.httk_rest),
         B_nonrest = log(Rest) - log(CLtot.httk_nonrest),
         C_rest = log(nonRest) - log(CLtot.httk_rest),
         C_nonrest = log(nonRest) - log(CLtot.httk_nonrest)) %>%
  select(Chemical, A_rest:C_nonrest) -> df_valcomp

df_valcomp %>% 
  column_to_rownames(var = "Chemical") %>%
  t() %>%
  as.matrix() %>%
  cor() %>%
  round(digits = 3) -> val_cormat

hclust(d = as.dist(1-abs(val_cormat)), method = "complete") %>% plot()

hclust(d = as.dist(1-abs(val_cormat)), method = "complete") -> val_hclust

corrplot::corrplot(abs(val_cormat))

#reorder the Chemical names prior to plotting
df_valcomp %>%
  mutate(Chemical = factor(x = Chemical,
                           levels = val_hclust$labels[val_hclust$order])) %>%
  pivot_longer(cols = A_rest:C_nonrest,
               names_to = "comparison", values_to = "logres") %>%
  mutate(comparison = case_when(
    comparison == "A_nonrest" ~ "free kelim vs httk nonrestrictive",
    comparison == "A_rest" ~ "free kelim vs httk restrictive",
    comparison == "B_nonrest" ~ "restrictive kelim vs httk nonrestrictive",
    comparison == "B_rest" ~ "restrictive kelim vs httk restrictive",
    comparison == "C_nonrest" ~ "nonrestrictive kelim vs httk nonrestrictive",
    comparison == "C_rest" ~ "nonrestrictive kelim vs httk restrictive",
  ),
  comparison = factor(comparison,
  levels = c("free kelim vs httk nonrestrictive",
             "free kelim vs httk restrictive",
             "restrictive kelim vs httk nonrestrictive",
             "restrictive kelim vs httk restrictive",
             "nonrestrictive kelim vs httk nonrestrictive",
             "nonrestrictive kelim vs httk restrictive"))) %>%
  ggplot(aes(x = comparison, y = Chemical, fill = logres)) +
  geom_tile() +
  scale_fill_gradient2(mid = "grey5", high = "magenta", low = "green",
                       name = "Total Clearance\nlog-scale difference") +
  coord_fixed(ratio = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, angle = 45),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank(),
        axis.title.x = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "Cltot_compare_heatmap_2024_0405.png"),
       height = 14,
       width = 8)

```{r dawson}
# try dawson
httk::load_dawson2021()
